{\bf Simulating a two dimensional particle in a square quantum box with CUDA}

### 2013-08-23 paper/paper.tex
Fixed the code
Improved the next probability function and added the implementation description in the paper about the maximum proba and the next probability

    \subsection{Brightness and Color mapping}
        \subsubsection{Brightness}
        The intensity of the colors denote the probability of finding the particle. Since
        probabilities are always between 0 and 1, converting them to intensities is just
        a matter of mapping the probabilities from $[0, 1]$ to $[0, 255]$.\\\\
        However $\epsilon$ is a small number, so the probability is always small; usually
        less than $0.1\%$. Therefore we need to find the highest probability in the space
        and remap from $[0, \text{max}]$ to $[0, 255]$. At first I have tried to brute force
        the problem exploiting my GPU using functions to find maximums using reduction. But
        the process of finding the highest probability of a particle with 10 energy levels
        inside a box divided into 262,144 boxes took around 28ms. Allowing me to get only
        35fps in the animation.\\
        A new solution needed to be adopted and I went over the equations again to try and
        find that maximum analytically. My results are:\\\\
        For $p(\alpha)$ to be maximal
        $$\frac{\epsilon}{L} - \frac{1}{n_{\alpha}\pi} \sin(\frac{n_{\alpha}\pi}{L}\epsilon)
        \cos{2\frac{n_{\alpha}\pi}{L}\alpha}$$
        needs to be maximal.  And this is achieved when $\cos{x} = -1$ which is possible
        since $0 \leq x \leq 2\pi$. Therefore the highest probability for one energy
        level is: 
        $$\frac{\epsilon}{L} + \frac{1}{n_{\alpha}\pi} \sin(\frac{n_{\alpha}\pi}{L}\epsilon)$$
        Following similar analysis we can deduce that the highest probability (denoted by
        $m$) is close to,
        but not exactly:
        $$m = m_x . m_y$$
        $$m_{\alpha} = \frac{\epsilon}{L} + \sum_i \left ( c^2_i \frac{1}{n_{\alpha_i}\pi}
        \sin\left(\frac{n_{\alpha_i}\pi}{L}\epsilon\right) \right )$$\\

@@

>>>>

    \subsection{Brightness and Color mapping}
        \subsubsection{Brightness}
        The intensity of the colors denote the probability of finding the particle. Since
        probabilities are always between 0 and 1, converting them to intensities is just
        a matter of mapping the probabilities from $[0, 1]$ to $[0, 255]$.\\\\
        However $\epsilon$ is a small number, so the probability is always small; usually
        less than $0.1\%$. Therefore we need to find the highest probability in the space
        and remap from $[0, \text{max}]$ to $[0, 255]$. At first I have tried to brute force
        the problem exploiting my GPU using functions to find maximums using reduction. But
        the process of finding the highest probability of a particle with 10 energy levels
        inside a box divided into 262,144 boxes took around 28ms. Allowing me to get only
        35fps in the animation.\\
        A new solution needed to be adopted and I went over the equations again to try and
        find that maximum analytically. My results are:\\\\
        For $p(\alpha)$ to be maximal
        $$\frac{\epsilon}{L} - \frac{1}{n_{\alpha}\pi} \sin\left(\frac{n_{\alpha}\pi}{L}\epsilon\right)
        \cos\left(2\frac{n_{\alpha}\pi}{L}\alpha\right)$$
        needs to be maximal.  And this is achieved when $\cos{x} = -1$ which is possible
        since $0 \leq x \leq 2\pi$. Therefore the highest probability for one energy
        level is: 
        $$\frac{\epsilon}{L} + \frac{1}{n_{\alpha}\pi} \sin(\frac{n_{\alpha}\pi}{L}\epsilon)$$
        Following similar analysis we can deduce that the highest probability (denoted by
        $m$) is close to,
        but not exactly:
        $$m = m_x . m_y$$
        $$m_{\alpha} = \frac{\epsilon}{L} + \sum_i \left ( c^2_i \frac{1}{n_{\alpha_i}\pi}
        \sin\left(\frac{n_{\alpha_i}\pi}{L}\epsilon\right) \right )$$\\\\
        \textbf{UPDATE} After implementing and testing the algorithm the uncertainty in the algorithm
        has proved to be overwhelming and the probabilities' range was no longer $[0, 1]$ but much smaller.
        Therefore a fallback was needed.\\\\
        The approach followed was a bruteforce solution where we compute the probabilities at each at each
        pixel and search for the largest. However to make the searching faster a reduction algorithm
        was adopted to speed it up. The algorithm creates a much smaller array in which it stores
        the maximum of a chunk of the original array. Then on the CPU we loop over the results (which are
        usually 32) and pick the maximum of these. The following illustration describes how the maximum
        reduction algorithm works
        \begin{figure}[hb]
            \centering
                \includegraphics[width=8cm]{graphics/max_reduction.png}
            \caption{Finding the maximum element in an array using reduction}
        \end{figure}
        The fact that this algorithm was used adds a restriction on the dimensions of the window. The height
        and the width of the window need to be a power of 2 for this algorithm to work effectively.
        No solution has been made up yet to generalize the dimensions of the window.

@@

        \subsection{Next set of probabilities}
        Every frame in the animation consists of a new set of probabilities close to the ones of the
        previous frame. A problem came up and that is how to generate new set of probabilities close
        to the ones in the previous frame to insure the smooth animation between the frames.\\\\
        The model adopted to insure the smoothness and the continuity in the probabilities consists
        of multiple sine waves where the period of one is 10 times more than its neighbor and so on.
        If plotted, the model looks like the following: (the sines were stretched vertically for a better
        visualization)
        \begin{figure}[hb]
            \centering
                \includegraphics[width=8cm]{graphics/sine_probabilities.png}
            \caption{Plotting
               $x(t) = 10\left| sin \left( \frac{0.1}{\pi}t \right) \right|$, 
               $y(t) = 10\left| sin \left( \frac{1}{\pi}t \right) \right|$, 
               $z(t) = 10\left| sin \left( \frac{10}{\pi}t \right) \right|$}
        \end{figure}
        After finding the values of each sine wave at a time $t$ we can normalize the answers so that
        their sum is $1$. Bellow is a table representing some values normalized.
        \begin{center}
            \begin{tabular}{| c | c | c | c |}
                \hline
                \textbf{Time} & \textbf{x} & \textbf{y} & \textbf{z} \\ \hline
                0.1 & 0.0091 & 0.0915 & 0.8994 \\
                0.2 & 0.0096 & 0.0957 & 0.8947 \\
                0.3 & 0.0104 & 0.1035 & 0.8861 \\
                0.4 & 0.0116 & 0.1159 & 0.8725 \\
                0.5 & 0.0136 & 0.1350 & 0.8515 \\
                0.6 & 0.0166 & 0.1648 & 0.8186 \\
                0.7 & 0.0215 & 0.2135 & 0.7649 \\
                0.8 & 0.0304 & 0.3006 & 0.6690 \\
                0.9 & 0.0490 & 0.4834 & 0.4675 \\
                1.0 & 0.0824 & 0.8102 & 0.1074 \\
                1.1 & 0.0479 & 0.4698 & 0.4822 \\
                1.2 & 0.0368 & 0.3590 & 0.6042 \\
                1.3 & 0.0322 & 0.3134 & 0.6544 \\
                1.4 & 0.0309 & 0.2987 & 0.6704 \\
                1.5 & 0.0317 & 0.3053 & 0.6630 \\
                1.6 & 0.0347 & 0.3324 & 0.6329 \\
                1.7 & 0.0405 & 0.3859 & 0.5736 \\
                1.8 & 0.0509 & 0.4818 & 0.4673 \\
                1.9 & 0.0701 & 0.6595 & 0.2704 \\
                2.0 & 0.0859 & 0.8022 & 0.1119 \\ \hline
            \end{tabular}
        \end{center}
        \begin{center}Table 1: Values from $x(t)$, $y(t)$ and $z(t)$ at different times\end{center}
        A general formula can be deduced to find the equation of the nth wave out of $N$ waves.
        \begin{align}
            P_{n/N}(t) = \left| sin\left( \frac{10^{2-(N-n)}}{\pi}t \right)\right|
        \end{align}
